---
title: "Homework_4"
output: html_document
date: "2023-12-02"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Homework 4:Structural Topic Modeling

```{r load packages}
library(knitr)
library(LDAvis)
library(readr)
library(tidyverse)
library(quanteda) # quantitative analysis of textual data  (https://quanteda.io/articles/quickstart.html)
library(quanteda.textplots) # complementary to quanteda, for visualization
library(tidytext) #conversion of text to and from tidy formats
library(cld3) # for language detection
library(servr) # will be used for visualization
library(topicmodels) # implementation of Latent Dirichlet Allocation, several topic models included
library(lda) # alternative to topicmodels, implementation of LDA only
library(stm) # for structural topic modeling
library(dplyr)
library(ggplot2)
library(textstem)
library(readr)

set.seed(5528) # set seed for reproductibility
PATH = "~/seminar5"
setwd(PATH)

```
### a) Load Data
```{r Load & inspect data}
data <- load(file.path(PATH, "data/hoc_speeches.Rdata"))
# Count of unique values for categorical columns
speeches %>% summarise(
  unique_names = n_distinct(name),
  unique_parties = n_distinct(party),
  unique_genders = n_distinct(gender)
)

```
### (b) Summary Statistics

```{r summary statistics}
# Basic summary of each column
summary(speeches)
```

```{r}
# Summary for categorical variables
speeches %>%
  group_by(gender) %>%
  summarise(count = n(), avg_ntoken = mean(ntoken))
```

```{r}
speeches %>%
  group_by(party) %>%
  summarise(count = n(), avg_ntoken = mean(ntoken))

```
### (c) Turn data into coprus object
```{r corpus}
corpus_speeches <- quanteda::corpus(speeches$speech)

summary(corpus_speeches)
```

### plotting more advanced statistics
```{r}
ggplot(speeches, aes(x = ntoken)) +
  geom_histogram(binwidth = 50, fill = "blue", color = "black") +
  labs(title = "Distribution of Speech Lengths", x = "Number of Tokens", y = "Frequency")

```

```{r}
# Boxplot for speech length by gender
ggplot(speeches, aes(x = gender, y = ntoken, fill = gender)) +
  geom_boxplot() +
  labs(title = "Speech Length by Gender", x = "Gender", y = "Number of Tokens")
```
```{r}
# Boxplot for speech length by party
ggplot(speeches, aes(x = party, y = ntoken, fill = party)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Speech Length by Party", x = "Party", y = "Number of Tokens")
```

### (d) Turn this corpus into a tokens object and then into a document-feature matrix
```{r}
# Tokenizing the corpus
tokens_speeches <- quanteda::tokens(corpus_speeches, remove_punct = TRUE, remove_numbers = TRUE)

# build the document-feature matrix
dfm_speeches <- dfm(tokens_speeches)

#plotfrequency
textplot_wordcloud(dfm_speeches, random_order = FALSE, rotation = 0.25, 
    color = RColorBrewer::brewer.pal(8, "Dark2"))

```

### stopwords and punctuation
```{r}
head(stopwords("en"), 50)

#stopwords
custom_stopwords <- read_csv(("data/list_stopwords.csv"), col_types = cols(X1 = col_skip()))
head(custom_stopwords,5)

custom_stopwords <- c(custom_stopwords$stop_words,stopwords("en"), 'also', 'may','can')

dfm_speechescorpus <- corpus_speeches %>%
                      quanteda::tokens(remove_punct = TRUE, remove_symbols = TRUE, remove_numbers = TRUE) %>%
                      quanteda::tokens_remove(custom_stopwords) %>%
                      quanteda::tokens_wordstem() %>%
                      quanteda::tokens_ngrams(n = c(1, 2)) %>%
                      dfm() %>%
                      dfm_tolower() %>%
                      dfm_trim(min_termfreq = 5, min_docfreq = 0.0025, docfreq_type = "prop")
                            
dfm_speechescorpus

```
```{r}
textplot_wordcloud(dfm_speechescorpus, random_order = FALSE, rotation = 0.25, 
    color = RColorBrewer::brewer.pal(8, "Dark2"),max_words =100,max_size = 3)
```

### (e & f) Run the STM
```{r tokens object}
set.seed(123)
stm_model = stm(dfm_speechescorpus, K = 20, max.em.its = 100, prevalence = ~ gender, data = speeches)

```

### (g) Plot STM
```{r plot}
plot(stm_model)
```

### (g) Examine top words from each topic
```{r examine}
# Examine the top words from each topic
topic_labels <- labelTopics(stm_model)

# Print the top words for each topic
print(topic_labels$prob)

```
```{r}
topic_descriptions <- c(
  "Parliamentary Questions",
  "Legislative Process",
  "Financial Institutions and Trade",
  "Parliamentary Membership and Speeches",
  "Parliamentary Terms and Rights",
  "Industry and Business",
  "Local Government and Education",
  "Public Concerns and Housing",
  "Local Investments and Environmental Issues",
  "International Relations and United Kingdom",
  "Governmental Statements and Operations",
  "Law, Justice, and Crime",
  "Government Finance and Taxation",
  "Political Parties and Elections",
  "Parliamentary Etiquette and Issues",
  "Employment and Labor Market",
  "Healthcare Services and Regional Governance",
  "Executive Governance and Negotiations",
  "Parliamentary Reporting and Committees",
  "Family, Gender, and Childhood"
)

```

### find top 3 documents associated with each topic
```{r, top 3}

K <- ncol(stm_model$theta)
cat("Number of topics:", K, "\n")

top_docs <- list()
for (k in 1:K) {
  topic_proportions <- stm_model$theta[, k]
  ordered_docs <- order(topic_proportions, decreasing = TRUE)[1:3]
  cat("Topic", k, "top docs:", ordered_docs, "\n")
  top_docs[[k]] <- ordered_docs
}

print(top_docs)
```

### Report top speeches for one selected topic
```{r}
topic_number <- 20

top_documents_indices <- order(stm_model$theta[, topic_number], decreasing = TRUE)[1:3]

top_speeches <- speeches$speech[top_documents_indices]

print(paste("Top speeches for topic", topic_number, ":"))
print(top_speeches)

```
### (h) Estimate the Effect of MP Gender on Topic Usage
(h) Use the estimateEffect and plot.estimateEffect functions in the stm package to estimate the effect of MP gender on topic usage. On which topics are women, on average, more active? 

### Answer:
Looking at the plots, we can see that the confidence interval for topic 18, Executive Governance and Negotiations, lies to the left side of the 0 line (a vertical line drawn at the 0 point on the x-axis, which represents no difference in topic prevalence between females and males) and does not cross it. This suggests that the effect is statistically significant and has stronger associations with males.

Further inspecting the plots, we can see that the dots to the right of the center (0) suggests that the topic is more prevalent in documents associated with female, although not statistically significant. The topics below show stronger prevalence with women and, on average, are more active with women:
"Legislative Process"                         "Financial Institutions and Trade"           
"Parliamentary Membership and Speeches"       "Parliamentary Terms and Rights"             
"Industry and Business"                       "Local Government and Education"             
"International Relations and United Kingdom"  "Political Parties and Elections"            
"Parliamentary Etiquette and Issues"          "Employment and Labor Market"                
"Healthcare Services and Regional Governance" "Parliamentary Reporting and Committees"    


```{r estimate effect}
effect_model = estimateEffect(formula = ~ gender, stmobj = stm_model, metadata = speeches)

dev.new(width = 17, height = 15)

plot.estimateEffect(effect_model, covariate = "gender", 
                    labeltype = "custom", 
                    custom.labels = stm_model$topics, 
                    xaxt = "n", 
                    cex.axis = 0.7, 
                    main = "Effect of Gender on Topic Prevalence")

```

```{r}

dev.new(width = 17, height = 15)

par(cex.axis = 0.7,  # Smaller axis text
    cex.lab = 0.7,   # Smaller x and y labels
    cex.main = 0.8)  # Slightly smaller main title

plot(effect_model, "gender", model=stm_model, method="difference",
     cov.value1="female", cov.value2="male")

```

```{r}
#topics with dots on the right of the zero line
selected_topics <- c(2, 3, 4, 5, 6, 7, 10, 14, 15, 16, 17, 19)

selected_labels <- topic_descriptions[selected_topics]

selected_labels
```